Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body

### 2019-05-22 draft_p2.tex
Reprise partielle de la presentation de GKTL

Let 
$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
refer formally to the probability of the trajectory 
$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$ of our dynamical system.
%
% $\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}} \right)$ 
%
By construction, the probability to obtain the same trajectory by resorting to the previous cloning procedure is given by
%
\begin{align}
\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)
\label{eq:Biased_Path_Approximation}
\end{align}
where the normalization factor is given by $Z(k,T_a)=\mathbb{E}_{0}\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right]$ with $\mathbb{E}_{0}$ being the expectation value with respect to $\mathbb{P}_{0}$.
%
% obvious? 
%and $\underset{N\rightarrow\infty}{\sim}$ meaning that this is true only asymptotically for large $N$.
%
%
%The typical error is of order $1/\sqrt{N}$ when evaluating averages over observables.
%
Eq.~\eqref{eq:Biased_Path_Approximation} relies on the mean-field approximation
\begin{equation}
R_{1}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_0}^{t_{1}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_1)= \mathbb{E}_{0}\left[e^{k\int_{t_0}^{t_{1}}A(\mathbf{X}(t))dt}\right]
\label{eq:Mean_Field_Approximation}
\end{equation}
which by induction (and assuming similar relations for the successive $R_i$'s) eventually leads  to~\cite{giardina_direct_2006,giardina_simulating_2011}
\begin{equation}
\prod_{i=1}^{T_{a}/\tau}R_{i}\underset{N\rightarrow\infty}{\sim} Z(k,T_a) =\mathbb{E}_{0}\left[e^{k\int_{0}^{T_a}A(\mathbf{X}(t))dt}\right].\label{eq:Estimate_Lambda}
\end{equation}
\EL{IS IT TRUE?}
The validity of this mean-field approximation and the typical relative error related to this approximation being of order $1/\sqrt{N}$ have been proven rigorously \cite{DelMoralBook,DelMoral2013} for a family of rare-event algorithms including the \ac{gktl} algorithm.
%
%
The killed trajectories, \emph{i.e.} with a weight smaller than $1$ \EL{true?}, are discarded from the statistics.
%
% no because the trajectories are cloned
%
% Starting from the final $N$ trajectories at time $T_{a}$, one goes backwards in time through the selection events attaching to each piece of trajectory its ancestor.
%
Eventually, an effective ensemble of $N$ trajectories for $0 \leq t \leq T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
All trajectories are solutions of the dynamical system since we have not modified the dynamics, but only sampled trajectories according to the distribution $\mathbb{P}_{k}$ rather than according to the distribution $\mathbb{P}_{0}$.
\EL{quid the small perturbation to separated the clones? comment required}
%
The sampled trajectories can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
%
More details concerning the implementation of the algorithm and the corresponding data analysis are provided in \cite{lestang:tel-01974316,brewer2018efficient}.

>>>>

Let 
$\mathbb{P}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)$ 
refer formally to the probability of the trajectory 
$\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}$ of our dynamical system.
%
% $\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}} = \left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}} \right)$ 
%
By construction, the probability to obtain the same trajectory by resorting to the previous cloning procedure is given by
%
\begin{align}
\mathbb{P}_{k}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right) &\underset{N\rightarrow\infty}{\sim} \frac{e^{k\int_{0}^{T_{a}}A(\mathbf{x}(t))dt}}{Z(k,T_a)}\mathbb{\mathbb{P}}_{0}\left(\left\{ \mathbf{X}(t)\right\} _{0\leq t\leq T_{a}}=\left\{ \mathbf{x}(t)\right\} _{0\leq t\leq T_{a}}\right)
\label{eq:Biased_Path_Approximation}
\end{align}
where the normalization factor is given by $Z(k,T_a)=\mathbb{E}_{0}\left[e^{k\int_{0}^{T_{a}}A(\mathbf{X}(t))dt}\right]$ with $\mathbb{E}_{0}$ being the expectation value with respect to $\mathbb{P}_{0}$.
%
% obvious? 
%and $\underset{N\rightarrow\infty}{\sim}$ meaning that this is true only asymptotically for large $N$.
%
%
%The typical error is of order $1/\sqrt{N}$ when evaluating averages over observables.
%
Eq.~\eqref{eq:Biased_Path_Approximation} relies on the mean-field approximation
\begin{equation}
R_{1}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_0}^{t_{1}}A(\mathbf{X}_{n}(t))dt}\underset{N\rightarrow\infty}{\sim} Z(k,t_1)= \mathbb{E}_{0}\left[e^{k\int_{t_0}^{t_{1}}A(\mathbf{X}(t))dt}\right]
\label{eq:Mean_Field_Approximation}
\end{equation}
The typical relative error related to this approximation can be shown of being of order $1/\sqrt{N}$\cite{DelMoralBook,DelMoral2013} for a family of rare-event algorithms including the \ac{gktl} algorithm.
which by induction (and assuming similar relations for the successive $R_i$'s) eventually leads  to~\cite{giardina_direct_2006,giardina_simulating_2011}
\begin{equation}
\prod_{i=1}^{T_{a}/\tau}R_{i}\underset{N\rightarrow\infty}{\sim} Z(k,T_a) =\mathbb{E}_{0}\left[e^{k\int_{0}^{T_a}A(\mathbf{X}(t))dt}\right].\label{eq:Estimate_Lambda}
\end{equation}
\EL{IS IT TRUE?}
The validity of this mean-field approximation and the 
%
%
The killed trajectories, \emph{i.e.} with a weight smaller than $1$ \EL{true?}, are discarded from the statistics.
%
% no because the trajectories are cloned
%
% Starting from the final $N$ trajectories at time $T_{a}$, one goes backwards in time through the selection events attaching to each piece of trajectory its ancestor.
%
Eventually, an effective ensemble of $N$ trajectories for $0 \leq t \leq T_{a}$ is obtained, distributed according to $\mathbb{P}_{k}$.
All trajectories are solutions of the dynamical system since we have not modified the dynamics, but only sampled trajectories according to the distribution $\mathbb{P}_{k}$ rather than according to the distribution $\mathbb{P}_{0}$.
\EL{quid the small perturbation to separated the clones? comment required}
%
The sampled trajectories can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
%
More details concerning the implementation of the algorithm and the corresponding data analysis are provided in \cite{lestang:tel-01974316,brewer2018efficient}.
